---
title: "Aggregate functions practice"
output: html_document
---

For these practice problems, we will use the gapminder data in the gapminder library (you must install it first). We also load the usual RSQLite package:
```{r}
library(gapminder)
library(RSQLite)
```

You can see the definitions of the columns by using the help() function:
```{r}
help(gapminder)
```

Then, we write the gapminder dataframe to a database table also named gapminder using the following code:
```{r}
con <- dbConnect(RSQLite::SQLite(), "")
dbWriteTable(con, "gapminder", gapminder)
```


1. Which year had the 3rd longest average life expectancy?
```{r}
query1 <- "select year, avg(lifeExp) as avgLife
from gapminder group by year order by avgLife desc"
dbGetQuery(con, query1)
```

2. Which continent had the largest average gdpPercap in 1992?
```{r}
query2 <- "select continent, avg(gdpPercap) as gdp from gapminder 
group by continent order by gdp desc"
dbGetQuery(con, query2)
```

3. How many African countries have data in the table for 2007?
```{r}
query3 <- "select continent, count(country) from gapminder where year = 2007 
group by continent"
dbGetQuery(con, query3)
```

4. What was the average gdpPercap for the continent with the lowest average life expectancy in 2007?
```{r}
query4 <- "select continent, avg(gdpPercap) as gdp, min(lifeExp) as life 
from gapminder where year = 2007 group by continent order by life "
dbGetQuery(con, query4)
```

5. Which country has the longest life expectancy, averaged over all years in the table?
```{r}
query5 <- "select country, avg(lifeExp) as life from gapminder 
group by country order by life desc"
dbGetQuery(con, query5)
```

6. How many countries have average life expectancies between 65 and 75, inclusive? (Average over all years in the table.)
```{r}
query6 <- "select country, avg(lifeExp) as life from gapminder 
group by country having life between 65 and 75"
dbGetQuery(con, query6)
```
```{r}
dbDisconnect(con)
```

















