---
title: "IntroSQL_practice"
output: html_document
---

We can turn any R dataframe into a database table. We'll do this mainly for practice problems. For these practice problems, we will use the flights data in the nycflights13 library (you must install it first).
```{r}
library(nycflights13)
```
The nycflights13 data contains five related dataframes: flights, airlines, airports, planes and weather. You can see the definitions of the columns by using the help() function, e.g.:
```{r}
help(flights)
```
Also, you can see how the five dataframes are related here: https://r4ds.had.co.nz/relational-data.html
We will talk more about these relations in the future.

We can put these dataframes into a database to use with SQL by using the following code:
```{r}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), "")
dbWriteTable(con, "flights", flights)
dbWriteTable(con,"airports", airports)
dbWriteTable(con,"airlines", airlines)
dbWriteTable(con,"planes", planes)
dbWriteTable(con,"weather", weather)
```
Note, this database does not exist physically on our computer like the Lahman Baseball Database we downloaded. Rather, it exists in our current R environment.

Answer the following questions, and enter your answer into the corresponding IntroSQL assignment on Mobius (accessed through Sakai).

1. How many tables are in the database we created? (Obviously you can just look at how we set it up, but get some practice using the R functions).
```{r}
dbListTables(con)
```

2. How many fields are in the weather table?
```{r}

length(dbListFields(con, "weather"))
```

3. How many fields are in the airlines table?
```{r}
length(dbListFields(con, "airlines"))
```

4. Write a query to retrieve all the rows for the month of January 2013 from the weather table. How many rows are returned?
```{r}
query4 = "SELECT * 
          FROM weather
          WHERE month = 1 AND year = 2013"
dbGetQuery(con, query4)

```

5. Write a query to retrieve all the rows for the 1st day of January, 2013 from the weather table. How many rows are returned?
```{r}
query5 = "SELECT *
          FROM weather
          WHERE month = 1 AND year = 2013 AND day = 1"
dbGetQuery(con, query5)

```