---
title: "JOIN_practice"
author: "AH"
date: "1/9/2021"
output: html_document
---

This assignment will give you practice using a JOIN in the context of a real-life data science problem (or, a very realistic simulation thereof).

Elkhart County (IN) has the highest percentage per capita of total workforce in the country employed in manufacturing. Close to 10,000 manufacturing jobs are left unfilled in Elkhart County because of a shortage of qualified workers. Every year, a Manufacturing Day event is held in conjunction with local businesses and schools to inform students about the career pathways available to them (https://heaindiana.org/student-pathways/pilot-programs/manufacturing-day/)

To assess the impact of the event, students are given a survey before and after Manufacturing Day with questions assessing their level of interest in manufacturing etc. 
Suppose students are asked to respond to the following item on both the before and after surveys:
I am interested in a career in manufacturing:
1=Strongly disagree
2=Somewhat disagree
3=Somewhat agree
4=Strongly agree

I have simulated some data to closely mimic actual data I have analyzed in the past. The R code below simulates the data and then writes it into a database connection named "con". The database has two tables, Pre_survey, which contains the survey responses before Manufacturing Day, and the Post_survey, which contains the responses after Manufacturing Day:
```{r}
library(RSQLite)
set.seed(3)
time1 <- sample(1:4, 100, prob = c(0.3, 0.3, 0.2, 0.2), replace = T)
time2 <- sample(1:4, 100, prob = c(0.2, 0.2, 0.3, 0.3), replace = T)
ID <- 1:100
df_pre <- data.frame(ID, time1)
df_pre_miss <- df_pre[-c(29, 70, 85), ]
df_post <- data.frame(ID, time2)
df_post_miss <- df_post[-c(17, 45, 56, 92), ]
con <- dbConnect(RSQLite::SQLite(), "")
dbWriteTable(con, "Pre_survey", df_pre_miss)
dbWriteTable(con,"Post_survey", df_post_miss)
dbListFields(con, "Pre_survey")
dbListFields(con, "Post_survey")
```
So, ID = student ID (uniquely identifies each student), time1 = response before Manufacturing Day, and time2 = response after Manufacturing Day.

1. Some students were absent the day of the pre-survey, and others were absent the day of the post-survey. Use a JOIN to create a query that includes ID and both the time1 and time2 variables. Furthermore, the query should include only students with responses to both the pre and post survey. How many records are returned?
```{r}
query1 <- "select Pre_survey.ID, time1, time2 from Pre_survey inner join Post_survey on Pre_survey.ID = Post_survey.ID" 
dbGetQuery(con, query1)
```

2. Now, we will compute a confidence interval to determine whether there was a significant change in the students’ attitude, on average. You can use the t.test() function in R. First, save the previous query as a data frame in R, and take look at the first few records:
```{r}
df1 <- dbGetQuery(con, query1)
head(df1)
```
Now that we have an R dataframe, we can use time2 and time1 as x and y, respectively, in the t.test() function. Report the lower bound of a two-sided 99% confidence interval for the true mean difference in attitudes between the two times. Round your answer to three decimal places. (You'll need to change a couple of the default arguments for the t.test() function, so you'll probably need to look at the help file).

```{r}
# help("t.test")
test_results <- t.test(df1$time2, df1$time1, conf.level = .99, paired = TRUE)
test_results

```

3. Does there appear to be a statistically significant difference between the two time points? (Use alpha = 0.01)
a. Yes
b. No

4. What is an appropriate interpretation of the confidence interval (CI)?
a. The CI is a plausible range for the true mean difference.
b. We are 99% confident the true mean difference lies in the CI.
c. There is a 99% probability the true mean difference lies in the CI.
d. If we sampled from the population many times and computed a 99% CI for each sample, then approximately 99% of the resulting CIs would contain the true mean difference.
e. All of the above are correct.
f. (a), (b), (d) are correct, but not (c)


Some notes (not needed for the actual assignment): If you are interested in being a data scientist, then you should be able to compute the above confidence interval and corresponding hypothesis test "by hand" (i.e. use R function to compute summary statistics, but know the formulas and be able to plug into them) as well as expound on the concepts in (3) and (4). If you are rusty with this, you should brush up on your 30600 notes, as these are fundamentals that you should have at your fingertips. On a different topic,in this activity we are treating an ordinal variable as continuous. This is done a lot, but there are alternate methods that preserve the ordinal nature of the data. Let me know if this sounds interesting.