---
title: "NULL Values practice"
output: html_document
---
Download the "portal_mammals" database here: https://github.com/weecology/portal-teachingdb/blob/master/portal_mammals.sqlite
Then, create a connection to the database:
```{r}
library(RSQLite)
con <- dbConnect(SQLite(), "portal_mammals.sqlite")
```
Examine the tables included in the database and look at the fields in each table in previous lessons. All subsquent items will refer to the "surveys" table in the mammals database.

1. How many rows are NULL for the hindfoot_length variable?
```{r}
query1 <- "select * from surveys where hindfoot_length is null"
dbGetQuery(con, query1)
```


2. How many rows are not NULL for the weight variable?
```{r}
query2 <- "select * from surveys where weight is not null"
dbGetQuery(con, query2)
```


3. How many rows are NULL for both the weight and hindfoot_length variables?
```{r}
query3 <- "select * from surveys where weight is null and hindfoot_length is null"
dbGetQuery(con, query3)
```


4. How many rows are NULL for at least one of weight and hindfoot_length?
```{r}
query4 <- "select * from surveys where weight is null or hindfoot_length is null"
dbGetQuery(con, query4)
```


5. How many rows are NULL for all three of weight, hindfoot_length, and species_id?
```{r}
query5 <- "select * from surveys where weight is null and species_id is null and 
hindfoot_length is null"
dbGetQuery(con, query5)
```

```{r}
dbDisconnect(con)
```














