---
title: "Primary Key Practice"
author: "AH"
date: "12/18/2020"
output: html_document
---
Load the RSQLite library:
```{r}
library(RSQLite)
```

1. Load the Lahman baseball database:
```{r}
con <- dbConnect(SQLite(), "lahmansbaseballdb.sqlite")
query1 <- "select * from Batting"
query2 <- "select count(distinct ID) from Batting"
dbGetQuery(con, query1)
dbGetQuery(con, query2)
```
Which column is the primary key for the Batting table?
a. stint
b. ID
c. yearID
d. lgID
e. There is no primary key



2. Connect to the NBA salary database:
```{r}
con <- dbConnect(SQLite(), "nba_salary.sqlite")
dbListTables(con)
dbListFields(con, "NBA_season1718_salary")
query3 <- "select * from NBA_season1718_salary"
dbGetQuery(con, query3)
query4 <- "select count(distinct Player) from NBA_season1718_salary"
dbGetQuery(con, query4)
```
Which column is the primary key for the NBA_season1718_salary table?
a. X1
b. Player
c. Tm
d. season17_18
e. There is no primary key.




3. Access the diamonds data in the tidyverse package, and save it as a database connection named con with a single table named diamonds:
```{r}
library(tidyverse)
con <- dbConnect(RSQLite::SQLite(), "diamonds")
dbWriteTable(con, "diamonds", diamonds)
```
What is the primary key for the diamonds table?
a. carat
b. clarity
c. table
d. primary
e. There is no primary key















