---
title: "WHERE clause Tutorial SOLUTION"
output: html_document
---
We can turn any R dataframe into a database table. We'll do this mainly for practice problems. For these practice problems, we will use the flights data in the nycflights13 library (you must install it first).
```{r}
library(nycflights13)
```

Then, we write the dataframe to a database table using the following code:
```{r}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), "flights")
dbWriteTable(con, "flights", flights)
```
As usual, the name of the connection is "con", and there is a single data table in the database, named "flights". 
`
1. How many rows are in the flights table?
```{r}
query1 <- "select * from flights"
dbGetQuery(con, query1)
```

2.There are three different airports in New York City, John F. Kennedy International, LaGuardia, and Newark. These are coded in the table as JFK, LGA, and EWR, respectively. How many records are for flights originating from JFK?
 
```{r}
query2 <- "select * from flights where origin = 'JFK'"
dbGetQuery(con, query2)
```


3. How many flights were greater than or equal to 1400 miles?
```{r}
query3 <- "select * from flights where distance >= 1400"
dbGetQuery(con, query3)
```


4. What was the distance flown by the longest flight originating from JFK?
```{r}
query4 <- "select * from flights where origin = 'JFK' order by distance desc"
dbGetQuery(con, query4)
```


5. How many flights originating from LaGuardia were greater than or equal to 1400 miles?
```{r}
query5 <- "select * from flights where origin = 'LGA' and distance >= 1400"
dbGetQuery(con, query5)
```


6. How many flights originating from LaGuardia and having Denver International Aiport (DEN) as the destination spent more than 250 minutes in the air?
```{r}
query6 <- "select * from flights where origin = 'LGA' and dest = 'DEN' and air_time > 250"
dbGetQuery(con, query6)
```


7. How many flights originating from LaGuardia and having Denver International Aiport (DEN) as the destination were in the air between 220 and 250 minutes (inclusive)?
```{r}
query7 <- "select * from flights where origin = 'LGA' and dest = 'DEN' and (air_time between 220 and 250)"
dbGetQuery(con, query7)
```


8. How many distinct destinations are in the data table?
```{r}
query8 <- "select count(distinct dest) from flights"
dbGetQuery(con, query8)
```


9. How many distinct destinations are in the data table for flights originating from JFK?
```{r}
query9 <- "select count(distinct dest) from flights where origin = 'JFK'"
dbGetQuery(con, query9)
```




```{r}
dbDisconnect(con) 
```